Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-30 paper/paper.tex
revise on section 3

At this point, we have obtained a linearized abstract abstract machine. If we imagine that the
classical AAM explores a graph of reachable states, then the linearized AAM flattens the
graph to a linear sequence of states.

@@

We have multiple choices of how to proceed:
\begin{itemize}
  \item By representing the continuations with monads\cite{filinski1994representing}, 
    we obtain the abstract definitional interpreters in monadic style which are similar to what 
    \citeauthor{darais2017abstracting} describe \cite{darais2017abstracting}.
  \item By representing the extended continuation-passing style with delimited control operators
    such as @shift@ and @reset@, we derive a new form of abstract interpreters in direct-style.
  \item In fact, we may also just use the same caching algorithm but with side effects such as
    assignments and mutations to update the cache, and then achieve the same definitional 
    interpreter with pushdown control-flows. \citeauthor{friedman-ai} in a tutorial on abstract 
    interpreters use this style to model caches and updates \cite{friedman-ai}.
    In this case, nondeterminism can be easily handled via @for@ comprehension. 
\end{itemize}

@@

\textit{Defunctionalization and Refunctionalization.}
\citeauthor{Reynolds:72} first shows defunctionalization as a program transformation 
technique that can be used to transform higher-order functions to first-order data types and 
its first-order dispatching function when writing definitional interpreters
\cite{Reynolds:72, Reynolds:HOSC98-revisited}. As generalized closure conversion,
defunctionalization is widely used in compilation and analysis \cite{pottier2006polymorphic, 
Eisenberg:2014:PFT:2633357.2633361, fourtounis2014modular, 10.1007/3-540-46425-5_4, consel1993tour}.
Refunctionalization~\cite{Danvy:2006:RW:2171265.2171268, Danvy:2001:DW:773184.773202} as a 
left inverse of defunctionalization, was firstly considered by \citeauthor{Danvy:2001:DW:773184.773202}.

\citeauthor{Ager:2003:FCE:888251.888254} revealed that defunctionalization and refunctionalization
can be used to show the functional correspondence between interpreters and abstract machines
\cite{Ager:2003:FCE:888251.888254}. A number of independently designed small-step semantic artifacts can 
be obtained by closure-converting, CPS-transforming, and then defunctionalizing their big-step counterparts, 
and vice versa by refunctionalizing \cite{DANVY2009534}. 
Such semantic artifacts include SECD machines \cite{10.1007/11431664_4, Ager:2003:FCE:888251.888254, DanvySECDJ},
CEK machines \cite{Ager:2003:FCE:888251.888254}, CLS machines  \cite{Ager:2003:FCE:888251.888254},
Categorical Abstract Machines \cite{Ager:2003:FCE:888251.888254}, lazy abstract machines \cite{AGER2004223}, 
monadic evaluators \cite{ager2005functional}, as well as languages with richer constructs
\cite{Danvy:2008:DIP:1411204.1411206, danvy2009towards, biernacka2009towards}.

>>>>

At this point, we have obtained a linearized abstract abstract machine. 
As \citeauthor{10.1007/3-540-15975-4_33} and \citeauthor{danvy2006a} 
showed, ``\textit{one programmer’s lazy-list constructors is another programmer’s success 
and failure continuations}'' \cite{10.1007/3-540-15975-4_33, danvy2006a}; the @step@ function
in AAM is the lazy-list constructor, and here we turn it into an explicit continuation.
Intuitively, if we imagine that the classical AAM explores a graph of reachable states, 
then the linearized AAM flattens the graph to a linear sequence of states.

@@

We have multiple choices of how to proceed:
\begin{itemize}
  \item By replacing the continuations with monads \cite{filinski1994representing}, 
    we obtain the abstract definitional interpreters in monadic style which are similar to what 
    \citeauthor{darais2017abstracting} describe \cite{darais2017abstracting}.
  \item By representing the extended continuation-passing style with delimited control operators
    such as @shift@ and @reset@, we derive a new form of abstract interpreters in direct-style.
  \item In fact, we may also just use the same caching algorithm but with side effects such as
    assignments and mutations to update the cache, and then achieve the same definitional 
    interpreter with pushdown control-flows. \citeauthor{friedman-ai} in a tutorial on abstract 
    interpreters use this style to model caches and updates \cite{friedman-ai}.
    In this case, nondeterminism can be easily handled via @for@ comprehension. 
\end{itemize}

@@

\textit{Defunctionalization and Refunctionalization.}
\citeauthor{Reynolds:72} first showed defunctionalization as a program transformation 
technique that can be used to transform higher-order functions to first-order data types and 
its first-order dispatching function when writing definitional interpreters
\cite{Reynolds:72, Reynolds:HOSC98-revisited}. As generalized closure conversion,
defunctionalization is widely used in compilation and analysis \cite{pottier2006polymorphic, 
Eisenberg:2014:PFT:2633357.2633361, fourtounis2014modular, 10.1007/3-540-46425-5_4, consel1993tour}.
Refunctionalization~\cite{Danvy:2006:RW:2171265.2171268, Danvy:2001:DW:773184.773202} as a 
left inverse of defunctionalization, was firstly considered by \citeauthor{Danvy:2001:DW:773184.773202}.

\citeauthor{Ager:2003:FCE:888251.888254} revealed that defunctionalization and refunctionalization
can be used to show the functional correspondence between interpreters and abstract machines
\cite{Ager:2003:FCE:888251.888254}. A number of independently designed small-step semantic artifacts can 
be obtained by closure-converting, CPS-transforming, and then defunctionalizing their big-step counterparts, 
and vice versa by refunctionalizing \cite{DANVY2009534}. 
Such semantic artifacts include SECD machines \cite{10.1007/11431664_4, Ager:2003:FCE:888251.888254, DanvySECDJ},
CEK machines \cite{Ager:2003:FCE:888251.888254}, CLS machines  \cite{Ager:2003:FCE:888251.888254},
Categorical Abstract Machines \cite{Ager:2003:FCE:888251.888254}, lazy abstract machines \cite{AGER2004223}, 
monadic evaluators \cite{ager2005functional}, as well as languages with richer constructs
\cite{Danvy:2008:DIP:1411204.1411206, danvy2009towards, biernacka2009towards}, etc.
